Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex
colors in Fig. 3
spacing in fig. 6
work on 1,2

\subsection{The Call Stack and Its Security}

{\em Functions} in high-level languages, and related abstractions such
as procedures, methods, etc., are units of computation
that call one another to modularly realize larger computations.
%
At a low level, each function activation manages its own data---local
variables, etc.---as well as information about the \emph{caller} to
which it must return.
%
The \emph{call stack} is the fundamental data structure used to
implement functions, aided by an ABI\bcp{will all CSF readers know
  what an ABI is?} that defines how registers are shared
between activations.

@@

Other mechanisms besides these should also be amenable to strong
formal guarantees.  In particular, Roessler and DeHon
\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety that offer
strong guarantees.  Their most realistic micro-policy, called Lazy
Tagging and Clearing (\(LTC\)), makes an interesting performance
trade-off: it allows function activations to improperly write into one
another's stack frames, but ensures that the owner of the corrupted
memory cannot access it afterward.

Under this policy, one function activation {\em can} corrupt another's
memory---just not in ways that affect observable behavior.
Therefore, \(LTC\) would not fulfill Georges et al.'s property
(adapted to the tagged setting).  But \(LTC\) does
arguably enforce stack safety.  A looser, more observational
definition of stack safety is needed to fit this situation.

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations from each other.
\bcp{Aren't other formal definitions based on the same intuition?}%
\sna{Two differences we're trying to get across. First, protecting
  functions vs. protecting stack frames: we bring registers into
  the story because it isn't just about the stack as a data structure.
  Second: describing security in terms of functions' high-level
  security goals (f is trying to extract a secret) rather than
  describing operationally what they can and can't do (as the CHERI
  folks do.)}%
\bcp{Both useful!  But neither comes across, at least not very clearly.}%
We use the tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
they are stated abstractly in hopes of applying
to other enforcement mechanisms.}

We find, via property-based random testing, that \(LTC\) is flawed in
a way that undermines both integrity and confidentiality; after
correcting this flaw, the repaired \(LTC\) satisfies all our
properties.  Moreover, we show that our properties can be extended to
describe calling conventions with features such as callee-save
registers, argument-passing on the stack, and tail-call
elimination. We modify \(LTC\) to protect these features and apply
random testing to validate this extended protection mechanism against
the extended properties.

\subsection {Towards Formal Properties for Stack Safety}

@@

Therefore, any low-level treatment of stack safety must begin by asking:
which high-level features does our system support, and what are their
security requirements? Different answers will beget different properties.
We identify and define five security requirements that a system might try
to enforce upon a function call: well-bracketed control flow, and the
integrity and confidentiality of each of the caller and the callee.
These are by no means exhaustive\bcp{That comment belongs in the
  future work
  section, not here}. We apply these first to a simple system
with very few features, then to a more realistic one supporting
tailcall-elimination, argument passing on the stack, and callee-saves
registers.

We show that our properties are useful for distinguishing correct enforcement
from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
testing tool for Coq. We generate random programs, using the features of our
more realistic system, and check that the stack safety {\em micro-policies}
of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
correctly detect the ones that attempt to violate one of our properties.

@@

\subsection{Contributions}

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}. \sna{Nowhere clearly enough. We need to make it clear in section 6
  that Nick and Andre are included there---only aiming at caller-side,
  not callee-side. The discussion of CHERI work should touch on this as well.}
The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We instantiate\bcp{extend??} these definitions in a realistic
setting with features such as\bcp{``such as''? are there others?}
argument passing on the stack, callee-saves registers, and tail-call
elimination.\bcp{How about: We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.}  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate a tag-based enforcement mechanism, \emph{Lazy
  Tagging and Clearing}, via property-based random testing, find that
it falls short, and propose and validate a fix.
\end{itemize}

In the next section, we give a brief overview of the structure of
properties.  In \cref{sec:example}, we walk through a function call in
a simple example machine and discuss how each of our properties
applies to it, informally. In the process we motivate the properties
from a security perspective.
%
In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
In \Cref{sec:enforcement} we describe the micro-policies that we test,
in \cref{sec:testing} the testing framework itself, and in
\cref{sec:relwork,sec:future} related and future work.

@@

\subsection{Threat Model and Limitations}
\apt{This could probably go better at end of next section.}

@@

Finally, we define high-level security properties---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts
that use variants\bcp{undefined},
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to internal implementation details.
We introduce these properties by example in \cref{sec:example} and
formally in \cref{sec:formal}. In the remainder
of this section we describe the underlying framework in more detail.

@@

\paragraph*{Machine model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer, and a standard ABI.
\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we distinguish the \emph{payload} part of
each state element---the part relevant to the behavior of the basic machine---from
any additional enforcement information (such as a tag).\bcp{Why?}
\sna{I think that the payload concept is too nitty-gritty for this section.
  The chief use of the payload is that it allows every value to be interpreted
  as an address. In the tag world, it happens to coincide with a natural instance
  of our variant-compatibility relation.}

@@

Security-relevant operations typically correspond to
\proposeadd{single?} underlying machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V architecture that we use in our
examples and tests,
calls and returns are conventionally performed using the {\tt jal}
(``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, but these
instructions might also be used for other things. An instantiation of
an operation
might require some external mechanism for distinguishing which instructions are
security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
which we use in our examples, is to add labels to selected instructions, but the theory does not
depend on this choice. One could instead recognize some operations dynamically. For instance,
if we make a call to untrusted (and unlabeled) code, we might decide to recognize
any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.

Whatever the mechanism for labeling transitions, we end up with an annotated version of the
machine transition function: we write \(\mach \xrightarrow{\bar{\psi},
  \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
an event, and \(\bar{\psi}\) is a list of security-relevant operations.
%
We then lift this into a transition between pairs of machine states and contexts,
by applying a set of operation-specific rules to transform each state and context to
a new context, in parallel with the ordinary transition on states.
The most important rules describe call and return operations.
A call pushes a new activation view\bcp{it was just called ``view'' above} onto the context stack and changes the class of the
caller's data to protect it from the new callee; a return reverses these steps.
Other operations can serve to signal how parts of the stack frame are being used to store
or share data, and their corresponding rules alter the classes of different
state elements accordingly.

@@

%% To characterize situations in which values should be kept secret or prevented
%% from influencing future execution, we borrow the idea of \emph{variant} states
%% from the theory of non-interference. Non-interference is a way of expressing
%% the flow of information through the system. If we begin with a state \(\mach\)
%% and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
%% then execute both \(\mach\) and \(\nach\), any differences in the outputs they
%% produce leaks information about \(\component\).

\paragraph*{Instantiating the Framework}

\proposecut{Our property framework has many parameters to support broad applicability.}
Conceptually, the following steps are needed to instantiate the framework to a specific machine
and code style: (i) define the base machine semantics; (ii) distinguish payload
from enforcement parts (if any) of the machine state; (iii) identify the set of
security-relevant operations and rules required by the code style; (iv) determine
how to label (or otherwise associate) machine instruction with security-relevant
operations as appropriate; (v) specify the form of observable events.

@@

\apt{Suggest moving threat model to here.}\bcp{Fine.}

@@

For a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
their values an only disagree on elements in \(\components\).
We say that the elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if they can be replaced by arbitrary other values without changing the
observable behavior of the machine. All other elements are \emph{relevant}.

(This story is slightly over-simplified. If an enforcement mechanism maintains
additional state associated with elements, such as tags, we don't want that
state to vary; only the payloads should. Formal definitions of variants and
relevance that incorporate this wrinkle are given in \cref{sec:props}.)

@@

To illustrate the issues, note that {\tt f} has also stored a 0 below the stack pointer.
%
\begin{figure}
    \centering
    \[
    \stackrel{\texttt{res}}{\figfourbox[0]}
    \stackrel{\texttt{sens}}{\figfourbox[0]}
    \stackrel{\texttt{sec}}{\figfourbox[5]}
    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
    %
    \[\overbrace{
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
    \]
    \[\large\Downarrow \hspace{1cm} \mathtt{a0} \leftbox[0] \hspace{1cm}
    \rightbox[7] \mathtt{a0} \hspace{1cm} \large\Downarrow\]
    \[
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
    \]
    \[\mathtt{a0} \leftbox[0] \hspace{1cm}
    \mathtt{a0} \rightbox[0]\]
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
and the register {\tt a0}. The execution from the right-hand state to its return
has changed the value at \(\SP - 4\), but that value matches that of the
left-hand variant at its return. Therefore, the change does not represent a leak.

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
The details of value structure depend on the specific machine being modelled;
intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, while the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag).
We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.
In the tagged setting, values are compatible if their tags are identical.
\apt{This is still not very convincing!
  empting to say that only the payloads can affect the step function, but that is
  too strong, right? (OK for tags, but not, e.g. capabilities?}
\sna{Yeah, I don't think we can say something universal about when values should be
  considered compatible.}

>>>>

%\subsection{The Call Stack and Its Security}

{\em Functions} in high-level languages, and related abstractions such
as procedures, methods, etc., are units of computation
that call one another to modularly realize larger computations.
%
At a low level, each function activation manages its own data---local
variables, etc.---as well as information about the \emph{caller} to
which it must return.
%
The \emph{call stack} is the fundamental data structure used to
implement functions, aided by an Application Binary Interface (ABI)
that defines how registers are shared between activations.

@@

Other mechanisms besides these should also be amenable to strong
formal guarantees.  In particular, Roessler and DeHon
\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety that offer
strong guarantees.  Their most realistic micro-policy, called Lazy
Tagging and Clearing (\(LTC\)), makes an interesting performance
trade-off: it allows function activations to improperly write into one
another's stack frames, but ensures that the owner of the corrupted
memory cannot access it afterward.
%
Under this policy, one function activation {\em can} corrupt another's
memory---just not in ways that affect observable behavior.
Therefore, \(LTC\) would not fulfill Georges et al.'s property
(adapted to the tagged setting).  But \(LTC\) does
arguably enforce stack safety.  A looser, more observational
definition of stack safety is needed to fit this situation.

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations (both registers and
stack frames) from each other.
%\bcp{Aren't other formal definitions based on the same intuition?}%
%% \sna{Two differences we're trying to get across. First, protecting
%%   functions vs. protecting stack frames: we bring registers into
%%   the story because it isn't just about the stack as a data structure.
%%   Second: describing security in terms of functions' high-level
%%   security goals (f is trying to extract a secret) rather than
%%   describing operationally what they can and can't do (as the CHERI
%%   folks do.)}%
%% \bcp{Both useful!  But neither comes across, at least not very clearly.}%
We use the tools of language-based
security~\cite{sabelfeld2003language}, treating function activations as
security principals.  We decompose stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
\proposecut{While our properties are motivated by the desire to specify \(LTC\) precisely,
they are stated abstractly in hopes of applying
to other enforcement mechanisms.}

%\subsection {Towards Formal Properties for Stack Safety}

@@

Therefore, any low-level treatment of stack safety must begin by asking:
which high-level features does our system support, and what are their
security requirements? Different answers will beget different properties.
We identify and define five security requirements that a system might try
to enforce upon a function call: well-bracketed control flow, and the
integrity and confidentiality of each of the caller and the callee.
%These are by no means exhaustive\bcp{That comment belongs in the
%  future work
%  section, not here}.
We apply these first to a simple system
with very few features, then to a more realistic one supporting
tail-call elimination, argument passing on the stack, and callee-save
registers.

We show that our properties are useful for distinguishing correct enforcement
from incorrect using QuickChick~\cite{Denes:VSL2014,Pierce:SF4}, a property-based
random testing tool for Coq.
We find that \(LTC\) is flawed in
a way that undermines both integrity and confidentiality; after
correcting this flaw, the repaired \(LTC\) satisfies all our
properties.  Moreover, we modify \(LTC\) to protect the features of
our more realistic system, and apply
random testing to validate this extended protection mechanism against
the extended properties.

%We generate random programs, using the features of our
%more realistic system, and check that the stack safety {\em micro-policies}
%of Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}
%correctly detect the ones that attempt to violate one of our properties.

@@

%\subsection{Contributions}

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}. \sna{Nowhere clearly enough. We need to make it clear in section 6
  that Nick and Andre are included there---only aiming at caller-side,
  not callee-side. The discussion of CHERI work should touch on this as well.}
The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item
  %We instantiate\bcp{extend??} these definitions in a realistic
%setting with features such as\bcp{``such as''? are there others?}
%argument passing on the stack, callee-saves registers, and tail-call
  %elimination.\bcp{How about:
 We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate a tag-based enforcement mechanism, \emph{Lazy
  Tagging and Clearing}, via property-based random testing, find that
it falls short, and propose and validate a fix.
\end{itemize}

In the next section, we give a brief overview of our framework and
assumptions. In \cref{sec:example}, we walk through a function call in
a simple example machine and discuss how each of our properties
applies to it, informally. In the process we motivate the properties
from a security perspective.
%
In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
In \Cref{sec:enforcement} we describe the micro-policies that we test,
in \cref{sec:testing} the testing framework itself, and in
\cref{sec:relwork,sec:future} related and future work.

@@

@@

To characterize situations in which values should be kept secret or prevented
from influencing future execution, we borrow the idea of \emph{variant} states
from the theory of non-interference.  For example, to show that certain locations
are kept secret, it suffices to vary their contents them and check that the observable
output does not change.

%Non-interference is a way of expressing
%the flow of information through the system. If we begin with a state \(\mach\)
%and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
%then execute both \(\mach\) and \(\nach\), any differences in the outputs they
%produce leaks information about \(\component\).

Finally, we define high-level security properties---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts
that use variants,
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to internal implementation details.
We introduce these properties by example in \cref{sec:example} and
formally in \cref{sec:formal}. In the remainder
of this section we describe the underlying framework in more detail.

@@

\paragraph*{Machine model}
We assume a conventional ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer. 
%, and a standard ABI.
%\sna{Not in our formalism (hardcoded +4 in the definition of WBCF notwithstanding.)
%  This assumption is made in our example and testing.}
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we keep the enforcement state unchanged when constructing variant states.
%% \sna{I think that the payload concept is too nitty-gritty for this section.
%%   The chief use of the payload is that it allows every value to be interpreted
%%   as an address. In the tag world, it happens to coincide with a natural instance
%%   of our variant-compatibility relation.}

@@

Security-relevant operations typically correspond to
single underlying machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V architecture that we use in our
examples and tests,
calls and returns are conventionally performed using the {\tt jal}
(``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, but these
instructions might also be used for other things. Thus we identify
security-relevant operations by labeling selected instructions. 
%% An instantiation of
%% an operation
%% might require some external mechanism for distinguishing which instructions are
%% security-relevant.\bcp{Huh?  What's an instantiation of an operation?} The most obvious technique,
%% which we use in our examples, is to add labels to selected instructions, but the theory does not
%% depend on this choice. One could instead recognize some operations dynamically. For instance,
%% if we make a call to untrusted (and unlabeled) code, we might decide to recognize
%% any jump back to trusted code as a return, to which \(\wbcf\) needs to apply.

We end up with an annotated version of the
machine transition function written \(\mach \xrightarrow{\bar{\psi},
  \obs} \mach'\), where \(\mach\) and \(\mach\) are machine states, \(\obs\) is
an event, and \(\bar{\psi}\) is a list of security-relevant operations.
%
We then lift this into a transition between pairs of machine states and contexts,
by applying a set of operation-specific rules to transform each state and context to
a new context, in parallel with the ordinary transition on states.
The most important rules describe call and return operations.
A call pushes a new view onto the context stack and changes the class of the
caller's data to protect it from the new callee; a return reverses these steps.
Other operations can serve to signal how parts of the stack frame are being used to store
or share data, and their corresponding rules alter the classes of different
state elements accordingly.

\apt{Still working from here down}

@@

\paragraph*{Instantiating the Framework}
Conceptually, the following steps are needed to instantiate the framework to a specific machine
and code style: (i) define the base machine semantics, including any hardware
security enforcement features; (ii) identify the set of
security-relevant operations and rules required by the code style; (iii) determine
how to label machine instruction with security-relevant
operations as appropriate; (iv) specify the form of observable events.

@@

\paragraph*{Threat Model and Limitations}

@@

For a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
their values an only disagree on elements in \(\components\).
We say that the elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if they can be replaced by arbitrary other values without changing the
observable behavior of the machine. All other elements are \emph{relevant}.\footnote{
This story is slightly over-simplified. If an enforcement mechanism maintains
additional state associated with elements, such as tags, we don't want that
state to vary. Formal definitions of variants and
relevance that incorporate this wrinkle are given in \cref{sec:props}.}

@@

To illustrate the issues, note that {\tt f} has also stored a 0 below the stack pointer.
%
\begin{figure}
    \centering
    \[
    \stackrel{\texttt{res}}{\figfourbox[0]}
    \stackrel{\texttt{sens}}{\figfourbox[0]}
    \stackrel{\texttt{sec}}{\figfourbox[5]}
    \stackrel{\texttt{ra}}{\figfourbox[0]\figfourbox[0]}\]
    %
    \[\overbrace{
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}}
    \]
    \[\large\Downarrow \hspace{1cm} \mathtt{a0} ~ \leftbox[0] \hspace{1cm}
    \rightbox[7] ~ \mathtt{a0} \hspace{1cm} \large\Downarrow\]
    \[
    \stackrel{\texttt{res}}{\leftbox[0]}
    \stackrel{\texttt{sens}}{\leftbox[0]}
    \stackrel{\texttt{sec}}{\leftbox[5]}
    \stackrel{\texttt{ra}}{\leftbox[0]\leftbox[0]}
    \hspace{1cm}
    %
    \stackrel{\texttt{res}}{\rightbox[2]}
    \stackrel{\texttt{sens}}{\rightbox[3]}
    \stackrel{\texttt{sec}}{\rightbox[4]}
    \stackrel{\texttt{ra}}{\rightbox[5]\rightbox[6]}
    \]
    \[\mathtt{a0} ~ \leftbox[0] \hspace{1cm}
    \rightbox[0] ~ \mathtt{a0}\]
  \caption{Return-time Confidentiality Violation}
  \label{fig:variant3}
\end{figure}
%
Now consider three elements: the address \(\SP - 4\), the address \(\SP + 12\),
and the register {\tt a0}. The execution from the right-hand state to its return
has changed the value at \(\SP - 4\), but that value matches that of the
left-hand variant at its return. Therefore, the change does not represent a leak.

@@

A {\em machine state} \(\mach \in \MACHS\) is a map from state elements to a set \(\mathcal{V}\) of
\emph{values}.
Each value \(v\) contains a \emph{payload} word, written \(|v|\).
We write \(\mach[\component]\) to denote the value of \(\mach\) at
\(\component\)  and \(\mach[v]\) as shorthand for \(\mach[|v|]\).
The details of value structure depend on the specific machine being modelled.
Intuitively, the payload represents the part of the value that is relevant to
the behavior of the basic machine, and it is the part that should be changed
when constructing variants (see below); the rest of the value may contain
information relevant to a hardware enforcement mechanism (such as a tag), which
should not be varied.
We assume a given \emph{compatibility} equivalence relation \(\sim\) on values,
and lift it element-wise to states.
In a tag-based enforcement scheme, values are compatible if their tags are identical.
%% \apt{This is still not very convincing!
%%   empting to say that only the payloads can affect the step function, but that is
%%   too strong, right? (OK for tags, but not, e.g. capabilities?}
%% \sna{Yeah, I don't think we can say something universal about when values should be
%%   considered compatible.}
